---
title: " Start with ilo package"
author: "David Bescond"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


ilo R tools
===========

This R package provides tools to manipulate 
[ILO Data from Statistics department database](http://www.ilo.org/ilostat/).

For contact information and source code, see the [github page](https://github.com/dbescond/ilo)

## Available tools

 * [Installation](#installation)  
 * [Download and Finding data](#search)  
 * [Get data](#get)
 * [Advanced manipulation](#advance)
 * [Save work](#save)
 * [Session](#session)


## <a name="installation"></a>Installation

Release version (not yet available):

```{r install, eval=FALSE}
# install.packages("ilo")
```

Development version:

```{r install2, eval=FALSE}
require(devtools)
install_github("dbescond/ilo")
```

Overall, the ilo package includes the following functions:

```{r, eval=TRUE}
require(ilo)
as.data.frame(ls("package:ilo")) 
```

check tutorials

```{r, eval=FALSE}
help_ilo()
```

## <a name="search"></a>Download and Finding data

Function `init_ilo()` download ilo Datasets and prepare setting. return dataframe of the available file `init_ilo(update)` force the update

```{r, eval=FALSE}
init_ilo()
```
setting are available in 3 languages, English, french and spanish `'en'`, `'fr'`, `'es'`} the default is `'en'`

```{r, eval=FALSE}
init_ilo(update, lang = 'fr')
```


Function `close_ilo()` close the clusters and `close_ilo(delete)` close and delete the entire repository
```{r, eval=FALSE}
close_ilo()
```


By default Database are store under your '.../'user'/AppData/Local/ILO/', lazyLoad rdb files are there. 

Then data are store by collection and freqency:

```{r, eval=FALSE}
ilo$segment 
```

get codelist availability
```{r, eval=FALSE}
names(ilo$code) %>% sort
```


## <a name="get"></a>Get and filter data 

Function `get_ilo()` collect data on one or several collection / freq, by default collection = 'YI', freq = 'A'

```{r, eval=FALSE}
# get monthly indicator (from STI collection) of Canada (geo = 'CAN')
get_ilo(collection = 'STI', freq = 'M', ref_area = 'CAN') 
```
```{r, eval=FALSE}
# get quarterly data (from ST collection) for unemployment rate (indicator = 'UNE_DEAP_RT')
get_ilo(collection = 'STI', freq = 'Q', ref_area = 'CAN', Query = "filter(indicator %in% 'UNE_DEAP_SEX_AGE_RT')")

```
Function `switch_ilo()` label data for one or several variables

```{r, eval=FALSE}
# label all variables and back in a wide style
 
Mydata <- get_ilo(collection = 'STI', freq = 'M', ref_area = 'ALB') 
Mydata %>% switch_ilo()

```
```{r, eval=FALSE}
# label only indicator
Mydata %>% switch_ilo(indicator)


```
Function `filter_ilo()` seach character string inside one variable
```{r, eval=FALSE}
Mydata <- get_ilo(collection = 'STI', freq = 'Q', ref_area = 'FRA') %>% switch_ilo(classif1, indicator) 
Mydata %>% filter_ilo(classif1 = 'Aggregate age bands')
Mydata %>% filter_ilo(indicator = 'Unemployment')

```
Check your data in with html tools (only for small dataset ...)
```{r, eval=FALSE}
get_ilo(collection = 'STI', freq = 'Q', ref_area = 'ALB') %>% 
  switch_ilo() %>% 
  view_ilo() 

```

## <a name="advance"></a> Advanced manipulation

This exercice show us how to make a special query ............. to do 

```{r, eval=FALSE}
LAST <- full_join(  
    get_ilo( query = "  select(ref_area, time, source) %>% 
                                    filter(str_sub(time,1,4) > '2013')", 
                       collection = 'STI', 
                       freq = 'Q', 
                       package= 'stringr') %>% 
        group_by(ref_area, time, source) %>%
        summarise(n = n()) %>% 
        ungroup() %>%
        spread(time, n) %>% filter(!substr(source,1,3) %in% 'GA:') 
        ,
# add the ref_area weight from collection = 'ILOEST' indicator %in% EAP_2EAP_NB   
      get_ilo( query = "    filter( as.character(time) %in% '2014', 
                                      indicator %in% 'EAP_2EAP_SEX_AGE_NB', 
                                      sex %in% 'SEX_T',
                          classif1 %in% 'AGE_5YRBANDS_TOTAL') %>%
                                  select(ref_area, obs_value)", 
                   collection = 'ILOEST', 
                   freq = 'A') %>% mutate(obs_value = round(obs_value, 0)) %>% 
      rename(EAP2014 = obs_value)
, by = 'ref_area') %>% 
select(ref_area, source, EAP2014, starts_with('20')) %>%
arrange(desc(EAP2014)) %>% # sort by the weight
mutate(iso3 = ref_area) %>% switch_ilo(ref_area)

```


## <a name="save"></a> Save your work


Save your data in xls (allowed multiple dataset)
```{r, eval=FALSE}
X <- get_ilo(collection = 'STI', freq = 'Q', ref_area = 'ALB')
Y <- get_ilo(ref_area = 'ALB') 

save_ilo(X = 'InSTI', Y = 'InYI')
save_ilo(X,Y)
```

Save your data in csv format
```{r, eval=FALSE}
X <- get_ilo(collection = 'STI', freq = 'Q', ref_area = 'ALB')
save_ilo(X, format = 'csv')
```

Save your data in csv ilo delete format
```{r, eval=FALSE}
X <- get_ilo(collection = 'STI', freq = 'Q', ref_area = 'ALB')
save_ilo(X, format = 'del')
```
Save your data in xls revision format (produce 2 sheets one for revision on for deletion)
```{r, eval=FALSE}
X <- get_ilo(collection = 'STI', freq = 'Q', ref_area = 'ALB')
save_ilo(X, format = 'rev')
```


## <a name="session"></a>Session info

This tutorial was created with

```{r sessioninfo, message=FALSE, warning=FALSE}
sessionInfo()
```
